---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
```

25 Slides

We would like your help to better understand factors that affect Job Slots product retention. Please use the data provided in Exhibit 1 and 2 to answer the following questions.																								
																								
We anticipate that it will take ~2-5 hours. You can analyze the data using whatever tools you feel will help you best extract insights.																								
Please present your responses to each question in a single PowerPoint deck (or PDF). In your write-up, include your assumptions and provide explanations for your answers. 																								
Additionally, please feel free to include any supporting materials you produced while addressing these question (scripts, model outputs, visualizations etc.). 																								
																								
PART A: Data wrangling in R (for reference, the orignial assignment for Part A asked for SQL. You are required to use R)	

```{r}
# we care about renewed_flag
perf = read_csv("PA446 Midterm Case Oct 27 - Ex 1 - slot_performance Table.csv")
locs = read_csv("PA446 Midterm Case Oct 27 - Ex 2 - location Table.csv")
```

1)	Total Contract Value is defined as the total amount that customers committed to spend. Write R tidyverse pipe that returns the Total Contract Value ('total_contract_value' field) for each state (name) by month (using 'start_date') in the provided datasets.	

```{r}
joined = perf %>% 
  left_join(locs, on=city_id) %>%
  mutate(start_date = mdy(start_date),
         start_month = month(start_date))

joined %>%
  group_by(state_name, start_month) %>%
  summarise(sum(total_contract_value))
```


```{r}
joined %>%
  group_by(state_name, month) %>%
  summarise
```
																								
2)	For all employers who purchased >1 product with [REDACTED], write a R tidyverse pipe to return the 'job_slots' and 'click_marketplace_value' values for the second transaction by employer.																							
																								
*Note: You only have to provide the R code needed to accomplish the questions above; output results are not required																								
																								
PART B: Metric Design																								
3)	What metrics would you propose to measure the quality of services [REDACTED] provided to our clients? How does performance vary in terms of:																							
	- job_slots?																							
	- total_contract_value?																							
	- click_marketplace_value?																							
																								
PART C: Retention Modeling and Analysis																								
4)	Which factors or combination of factors best correlate with an employer's likelihood to retain (i.e. renewed_flag = 1)? And how well does your chosen method correlate with retention? Please list any assumptions you made and explain why you chose your methodology.																							
																								
5)	Based on your analysis, what modifications would you recommend we make to our ad platform algorithm to improve retention?

